---
title: Phenology masks with temporal buffer
author: Guido Fioravanti
date: "`r Sys.Date()`"
---

This document illustrates the 36 phenology masks for masking fAPAR when crop is out of the growing seasons.

For each dekad (the number of each dekad in the title), two maps are shown:

- the phenology mask 

- the phenology mask with a temporal buffer of 1 dekad (as suggested by Willem): in this case, the beginning (the end) of the growing season is anticipated (psoticipated) of 1 daked

In the following maps NA values correspond to cells with "no crop", e.g.: urban areas, forests and so on..

A "difference map" highlights the effect of the application of a temporal buffer (in this case, the black dots correspond to crop cell where the crop is "active" in the map with the temporal buffer and "no active" in the map without the temporal buffer). As expected, the "buffered" maps exhibit a larger number of "active crop" cells.

```{r intro,include=FALSE,warning=FALSE,message=FALSE,echo=FALSE,error=FALSE}
rm(list=objects())
library("terra")
library("purrr")
library("rworldmap")
library("sf")
library("dplyr")
library("ggplot2")
library("ggspatial")
library("patchwork")
library("tidyterra")
library("scico")
#library("furrr")

#plan(strategy=multicore,workers=2)

rworldmap::countriesCoarse->mondo
cleangeo::clgeo_Clean(mondo)->mondo
st_as_sf(mondo) %>%
  filter(GEOUNIT=="Spain")->mondo

vect(mondo)->vmondo

list.files(pattern="^.+tif")->ffile


numeroFile<-length(ffile)
```


```{r,include=TRUE,warning=FALSE,message=FALSE,echo=FALSE,error=FALSE,out.width="100%"}
purrr::map(1:numeroFile,.f=function(.qualeFile){ #
  
  rast(ffile[.qualeFile])->grid_not_masked
  crop(grid_not_masked,mondo)->grid_not_masked
  
  rast(glue::glue("./ris_right/buffered_{ffile[.qualeFile]}"))->grid_masked 
  names(grid_masked)<-paste0("buffered_",names(grid_masked))
  
  grid_not_masked-grid_masked->differenza
  names(differenza)<-"difference"
  


  #classify
  classify(grid_masked,rcl=matrix(data=c(-1,0,0,0,1,1),nrow=2,byrow=TRUE))->grid_masked
  levels(grid_masked)<-c("no active crop","active crop")

  classify(grid_not_masked,rcl=matrix(data=c(-1,0,0,0,1,1),nrow=2,byrow=TRUE))->grid_not_masked
  levels(grid_not_masked)<-c("no active crop","active crop")


  
  stringr::str_remove(ffile[.qualeFile],"01dd.tif")->titolo
  ggplot()+
    ggspatial::layer_spatial(grid_not_masked)+
    ggspatial::layer_spatial(mondo,fill="transparent",color="black")+
    ggtitle(paste0(titolo," (no buffer)"))+
    scale_fill_scico_d(na.value="transparent",palette="bamako")+
    theme_bw()+
    theme(panel.grid=element_blank())->grafico1
  
  ggplot()+
    ggspatial::layer_spatial(grid_masked)+
    ggspatial::layer_spatial(mondo,fill="transparent",color="black")+
    ggtitle(paste0(titolo," (with buffer)"))+    
    scale_fill_scico_d(na.value="transparent",palette="bamako")+
    theme_bw()+
    theme(panel.grid=element_blank())->grafico2
  
  ggplot()+
    ggspatial::layer_spatial(differenza)+
    ggspatial::layer_spatial(mondo,fill="transparent",color="black")+
    ggtitle(paste0("Difference between the two maps"))+
    scale_fill_scico(na.value="transparent",palette="bamako",guide="none")+
    theme_bw()+
    theme(panel.grid=element_blank())->graficoDiff
  
    print((grafico1+grafico2+plot_layout(guides="collect",nrow=2,ncol = 2))/(graficoDiff+plot_layout(design="1#")))

})->listaOut1
```

